---
title: "Thesis Analysis Code"
author: "Moody Billah"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(DBI)
library(RMySQL)
library(sqldf)
library(tidyverse)
library(car)
library(Rmisc)
```

Reading and cleaning web scraped data with 719 observations

```{r}
con = dbConnect(MySQL(), user='root', password='root', dbname='craigslist_web_scraping', host='localhost')
query = "select rent, bedrooms, sq_feet, page_date from listings l join webpages w on l.page_ID = w.page_ID where rent > 0"
craigslist_df = dbGetQuery(con, query)

craigslist_df = mutate(craigslist_df, bedrooms = ifelse(bedrooms < 0, 0, bedrooms))
craigslist_df = mutate(craigslist_df, bedrooms = ifelse(bedrooms > 5, 5, bedrooms))
craigslist_df = mutate(craigslist_df, sq_feet = ifelse(sq_feet < 0, NA, sq_feet))
craigslist_df$bedrooms = as.factor(craigslist_df$bedrooms)

dim(craigslist_df)
head(craigslist_df)
```

Creating pandemic variable with the cutoff at date at 2020-03-11

```{r}
craigslist_df$page_date = as.Date(craigslist_df$page_date, format="%Y-%m-%d")
craigslist_df$pandemic = as.factor(ifelse(craigslist_df$page_date >= "2020-03-11", 1, 0))
```

Overall distribution

```{r}
ggplot(craigslist_df, aes(x=rent)) + geom_histogram(binwidth=100) + ggtitle("Distribution of Overall Rent")
```

Distributions by pandemic

```{r}
ggplot(craigslist_df, aes(x=rent, fill=pandemic)) + geom_histogram(binwidth=100) + ggtitle("Distribution of Rent by Pandemic")
```

Significant difference pre and post pandemic (ANOVA assumptions met)

```{r}
aov_pandemic = aov(rent ~ pandemic, craigslist_df)
summary(aov_pandemic)
```

```{r}
plot(aov_pandemic, which=c(2,3))
```


Distribution by bedrooms

```{r}
ggplot(craigslist_df, aes(x=rent, fill=bedrooms)) + geom_histogram(binwidth=100) + ggtitle("Distribution of Rent by Bedrooms")
```

Significant difference between bedrooms (ANOVA assumptions met)

```{r}
aov_bedrooms = aov(rent ~ bedrooms, craigslist_df)
summary(aov_bedrooms)
```

```{r}
plot(aov_bedrooms, which=c(2,3))
```

Distribution by pandemic and bedrooms

```{r}
pre_pandemic = filter(craigslist_df, pandemic == 0)
post_pandemic = filter(craigslist_df, pandemic == 1)

ggplot(pre_pandemic, aes(x=rent, fill=bedrooms)) + geom_histogram(binwidth=100) + ggtitle("Distribution of Rent by Bedrooms Pre Pandemic")
```

```{r}
ggplot(post_pandemic, aes(x=rent, fill=bedrooms)) + geom_histogram(binwidth=100) + ggtitle("Distribution of Rent by Bedrooms Post Pandemic")
```

No significant interaction between pandemic and bedrooms (ANOVA assumptions met)

```{r}
aov_pan_bed = aov(rent ~ pandemic*bedrooms, craigslist_df)
summary(aov_pan_bed)
```

```{r}
plot(aov_pan_bed, which=c(2,3))
```


Square feet and bedrooms have high multicollinearity so square feet will be ignored from analysis (linear regression assumptions met)

```{r}
lm_sq_bed = lm(rent ~ sq_feet + bedrooms + pandemic, craigslist_df)
summary(lm_sq_bed)
vif(lm_sq_bed)
```

```{r}
plot(lm_sq_bed, which=c(1,2))
```

Generating ACS bedroom data pre pandemic

```{r}
set.seed(100)
acs_beds_prepan = sample(c(0,1,2,3,4,5), length(pre_pandemic$bedrooms), replace=T, 
                         prob=c(0.071,0.191,0.299,0.283,0.123,0.033))
```

Generating ACS bedroom data post pandemic

```{r}
set.seed(100)
acs_beds_postpan = sample(c(0,1,2,3,4,5), length(post_pandemic$bedrooms), replace=T, 
                         prob=c(0.072,0.192,0.298,0.281,0.123,0.034))
```

Creating function to generate ACS rent data

```{r}
acs_generate_rent = function(rindex, max_rent, n){
  acs_rent = rep(0, n)
  for (i in 1:n){
  acs_rent[i] = case_when(
    rindex[i] == 1 ~ runif(1, 0, 499),
    rindex[i] == 2 ~ runif(1, 500, 999),
    rindex[i] == 3 ~ runif(1, 1000, 1499),
    rindex[i] == 4 ~ runif(1, 1500, 1999),
    rindex[i] == 5 ~ runif(1, 2000, 2499),
    rindex[i] == 6 ~ runif(1, 2500, 2999),
    rindex[i] == 7 ~ runif(1, 3000, max_rent)
    )
  }
  return(acs_rent)
}
```

Generating ACS rent data pre pandemic

```{r}
set.seed(100)
rindex_prepan = sample(c(1,2,3,4,5,6,7), length(pre_pandemic$bedrooms), replace=T, 
                       prob=c(0.046,0.147,0.334,0.243,0.122,0.057,0.052))

acs_rent_prepan = acs_generate_rent(rindex_prepan, max(pre_pandemic$rent), length(pre_pandemic$rent))
```

Generating ACS rent data post pandemic

```{r}
set.seed(100)
rindex_postpan = sample(c(1,2,3,4,5,6,7), length(post_pandemic$bedrooms), replace=T, 
                        prob=c(0.045,0.127,0.311,0.253,0.135,0.065,0.064))

acs_rent_postpan = acs_generate_rent(rindex_postpan, max(post_pandemic$rent), length(post_pandemic$rent))
```

Distribution of ACS rent 

```{r}
acs_df = rbind(data.frame(rent=acs_rent_prepan, pandemic=as.factor(0)), 
               data.frame(rent=acs_rent_postpan, pandemic=as.factor(1)))

ggplot(acs_df, aes(x=rent)) + geom_histogram(binwidth=100) + ggtitle("Distribution of Overall ACS Rent")
```

Distribution of ACS rent by pandemic

```{r}
ggplot(acs_df, aes(x=rent, fill=pandemic)) + geom_histogram(binwidth=100) + ggtitle("Distribution of ACS Rent by Pandemic")
```

No significant difference on ACS pre and post pandemic (normality assumption skewed right)

```{r}
aov_acs_pandemic = aov(rent ~ pandemic, acs_df)
summary(aov_acs_pandemic)
```

```{r}
plot(aov_acs_pandemic, which=c(2,3))
```

Significant difference between ACS and Craigslist rent overall

```{r}
t.test(craigslist_df$rent, acs_df$rent)
```

```{r}
ggplot() + geom_histogram(data=craigslist_df, aes(x=rent, fill="Craigslist"), binwidth=100, alpha=0.5) + 
  geom_histogram(data=acs_df, aes(x=rent, fill="ACS"), binwidth=100, alpha=0.5) + 
  ggtitle("Distribution of Overall ACS vs Craigslist Rent") + scale_fill_discrete(name="source")
```

Significant difference between ACS and Craigslist rent pre pandemic

```{r}
t.test(pre_pandemic$rent, acs_rent_prepan)
```

```{r}
prepan_acs = filter(acs_df, pandemic == 0)

ggplot() + geom_histogram(data=pre_pandemic, aes(x=rent, fill="Craigslist"), binwidth=100, alpha=0.5) + 
  geom_histogram(data=prepan_acs, aes(x=rent, fill="ACS"), binwidth=100, alpha=0.5) + 
  ggtitle("Distribution of Pre Pandemic ACS vs Craigslist Rent") + scale_fill_discrete(name="source")
```

Significant difference between ACS and Craigslist rent post pandemic

```{r}
t.test(post_pandemic$rent, acs_rent_postpan)
```

```{r}
postpan_acs = filter(acs_df, pandemic == 1)

ggplot() + geom_histogram(data=post_pandemic, aes(x=rent, fill="Craigslist"), binwidth=100, alpha=0.5) + 
  geom_histogram(data=postpan_acs, aes(x=rent, fill="ACS"), binwidth=100, alpha=0.5) + 
  ggtitle("Distribution of Post Pandemic ACS vs Craigslist Rent") + scale_fill_discrete(name="source")
```

Getting 95% confidence interval of Craigslist rent by bedroom

```{r}
beds = sort(unique(craigslist_df$bedrooms))
upper = mean = lower = rep(0, length(beds))

for (i in 1:length(beds)){
  bed_rent = craigslist_df %>% filter(bedrooms == i-1) %>% select(rent)
  upper[i] = CI(bed_rent$rent)[1]
  mean[i] = CI(bed_rent$rent)[2]
  lower[i] = CI(bed_rent$rent)[3]
}

beds_confint = cbind.data.frame(beds, lower, mean, upper)
beds_confint
```

Getting the percentiles from overall rent by bedroom using 95% confidence interval bounds

```{r}
beds_confint$lower_percent = ecdf(craigslist_df$rent)(beds_confint$lower)
beds_confint$upper_percent = ecdf(craigslist_df$rent)(beds_confint$upper)
beds_confint
```

Creating function to associated ACS rent with bedrooms using Gibbs sampling

```{r}
lower_pct = beds_confint$lower_percent
upper_pct = beds_confint$upper_percent

gibbs_sampler = function(acs_beds, acs_rent, lower_pct, upper_pct){
  acs_rentbed = rep(0, length(acs_beds))
  t = 1
  while (t <= 100){
    for (i in 1:length(acs_beds)){
      beds_val = acs_beds[i]
      lower_cut = quantile(acs_rent, lower_pct[beds_val+1])
      upper_cut = quantile(acs_rent, upper_pct[beds_val+1])
      cut_df = data.frame("rent" = acs_rent) %>% filter(rent >= lower_cut, rent <= upper_cut)
      acs_rentbed[i] = sample(cut_df$rent, 1)
    }
    t = t + 1
  }
  return(acs_rentbed)
}
```

Simulating Gibbs sample for pre pandemic ACS rent

```{r}
set.seed(100)
acs_rentbed_prepan = gibbs_sampler(acs_beds_prepan, acs_rent_prepan, lower_pct, upper_pct)
```

Simulating Gibbs sample for post pandemic ACS rent

```{r}
set.seed(100)
acs_rentbed_postpan = gibbs_sampler(acs_beds_postpan, acs_rent_postpan, lower_pct, upper_pct)
```

Distribution of ACS rent by bedroom pre pandemic

```{r}
acs_sim_prepan = data.frame("bedrooms" = as.factor(acs_beds_prepan), "rent" = acs_rentbed_prepan)

ggplot(acs_sim_prepan, aes(x=rent, fill=bedrooms)) + geom_histogram(binwidth=300) + ggtitle("Distribution of ACS Rent by Bedrooms Pre Pandemic")
```

Distribution of ACS rent by bedroom post pandemic

```{r}
acs_sim_postpan = data.frame("bedrooms" = as.factor(acs_beds_postpan), "rent" = acs_rentbed_postpan)

ggplot(acs_sim_postpan, aes(x=rent, fill=bedrooms)) + geom_histogram(binwidth=300) + ggtitle("Distribution of ACS Rent by Bedrooms Post Pandemic")
```

Distribution of ACS rent by bedroom overall

```{r}
acs_sim_df = rbind.data.frame(acs_sim_prepan, acs_sim_postpan)

ggplot(acs_sim_df, aes(x=rent, fill=bedrooms)) + geom_histogram(binwidth=300) + ggtitle("Distribution of ACS Rent by Bedrooms Overall")
```

Significant difference between Craigslist and ACS for 0 bedroom

```{r}
cl_b0 = craigslist_df %>% filter(bedrooms == 0) %>% select(rent)
acs_b0 = acs_sim_df %>% filter(bedrooms == 0) %>% select(rent)

t.test(cl_b0$rent, acs_b0$rent)
```

```{r}
ggplot() + geom_histogram(data=cl_b0, aes(x=rent, fill="Craigslist"), binwidth=100, alpha=0.5) + 
  geom_histogram(data=acs_b0, aes(x=rent, fill="ACS"), binwidth=100, alpha=0.5) + 
  ggtitle("Distribution of 0 Bedroom ACS vs Craigslist Rent") + scale_fill_discrete(name="source")
```

Significant difference between Craigslist and ACS for 1 bedroom

```{r}
cl_b1 = craigslist_df %>% filter(bedrooms == 1) %>% select(rent)
acs_b1 = acs_sim_df %>% filter(bedrooms == 1) %>% select(rent)

t.test(cl_b1$rent, acs_b1$rent)
```

```{r}
ggplot() + geom_histogram(data=cl_b1, aes(x=rent, fill="Craigslist"), binwidth=100, alpha=0.5) + 
  geom_histogram(data=acs_b1, aes(x=rent, fill="ACS"), binwidth=100, alpha=0.5) + 
  ggtitle("Distribution of 1 Bedroom ACS vs Craigslist Rent") + scale_fill_discrete(name="source")
```

Significant difference between Craigslist and ACS for 2 bedroom

```{r}
cl_b2 = craigslist_df %>% filter(bedrooms == 2) %>% select(rent)
acs_b2 = acs_sim_df %>% filter(bedrooms == 2) %>% select(rent)

t.test(cl_b2$rent, acs_b2$rent)
```

```{r}
ggplot() + geom_histogram(data=cl_b2, aes(x=rent, fill="Craigslist"), binwidth=100, alpha=0.5) + 
  geom_histogram(data=acs_b2, aes(x=rent, fill="ACS"), binwidth=100, alpha=0.5) + 
  ggtitle("Distribution of 2 Bedroom ACS vs Craigslist Rent") + scale_fill_discrete(name="source")
```

Significant difference between Craigslist and ACS for 3 bedroom

```{r}
cl_b3 = craigslist_df %>% filter(bedrooms == 3) %>% select(rent)
acs_b3 = acs_sim_df %>% filter(bedrooms == 3) %>% select(rent)

t.test(cl_b3$rent, acs_b3$rent)
```

```{r}
ggplot() + geom_histogram(data=cl_b3, aes(x=rent, fill="Craigslist"), binwidth=100, alpha=0.5) + 
  geom_histogram(data=acs_b3, aes(x=rent, fill="ACS"), binwidth=100, alpha=0.5) + 
  ggtitle("Distribution of 3 Bedroom ACS vs Craigslist Rent") + scale_fill_discrete(name="source")
```

Significant difference between Craigslist and ACS for 4 bedroom

```{r}
cl_b4 = craigslist_df %>% filter(bedrooms == 4) %>% select(rent)
acs_b4 = acs_sim_df %>% filter(bedrooms == 4) %>% select(rent)

t.test(cl_b4$rent, acs_b4$rent)
```

```{r}
ggplot() + geom_histogram(data=cl_b4, aes(x=rent, fill="Craigslist"), binwidth=100, alpha=0.5) + 
  geom_histogram(data=acs_b4, aes(x=rent, fill="ACS"), binwidth=100, alpha=0.5) + 
  ggtitle("Distribution of 4 Bedroom ACS vs Craigslist Rent") + scale_fill_discrete(name="source")
```

Not enough data points to investigate difference between Craigslist and ACS for 5 bedroom

